---
title: "WGCNA_prep_subset"
output: html_document
date: '2022-10-11'
---

##Objectives: 
1. Identify appropriate beta parameter based on scale free topology 
1a. I am comparing a range of parameters for a signed network (network )
2. Create adjacency table 
2a. Compare the different adjacency tables produced by correlation tests (bi weight vs pearson). I can use quantiles of the adjacency matrix to compare between them. 
3. Transform adjacency table to topological overlap matrix 
4. Determine Modules
4a. Can test different tree cut parameters? 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(WGCNA)
load("combat_adj_91_samples.Rdata")
combat.adj = t(combat.adj)
```

##Network Definitions 
Density = mean(k)/n-1 which is approximately mean(k)/n; n is the number of nodes and k = (k1,...,kn) denotes the vector of node connectivities. A density close to 1 indicates that all individuals strongly like each other, while a density of 0.5 suggests the presence of more ambiguous relationships. Module detection (and clustering) methods aim to find subnetworks with high density

Centralization is approximately equal to max(k)/n - Density. The centralization is 1 for a network with star topology; by contrast, it is 0 for a network where each node has the same connectivity. 

Heterogeneity is based on the variance of the connectivity across the nodes. It is defined as the sqrt(var(k))/mean(k) 


#Pearson Correlation
"powerEstimate = estimate of an appropriate soft-thresholding power: the lowest power which the scale free topology fit R^2 exceeds RsquaredCut. If R^2 is below RsquaredCut for all powers, NA is returned 

RsquaredCut is the desired minimum scale free topology fitting index R^2; the default value in the function for pickSoftThreshold is 0.85. 

pst[["powerEstimate"]] = 11; RsquaredCut = 0.90

```{r}
powers=seq(from=2,to=20,by=1) 
sft_pearson=pickSoftThreshold(combat.adj,RsquaredCut = 0.90, powerVector=powers, networkType ="signed", moreNetworkConcepts=T)
sft[["powerEstimate"]]
pst = sft_pearson[["fitIndices"]]

cex1=1.1
cex.axis=1.5
cex.lab=1.5
cex.main=1.5 
par(mfrow=c(2,2));
plot(powers,-sign(pst[,3])*pst[,2],type="n",
     xlab="Soft Threshold",
     ylab="SFT,signed Rˆ2",cex.axis=cex.axis, cex.main=cex.main, cex.lab=cex.lab, main="Scale Free Fit Index Rˆ2");
text(powers,-sign(pst[,3])*pst[,2], labels=powers,cex=cex1,col="red")
# this line corresponds to using an Rˆ2 cut-off of h 
abline(h=0.95,col="red")
plot(powers,pst$Density,
     type="n", xlab="Soft Threshold",ylab="Density", cex.axis=cex.axis, cex.main=cex.main,cex.lab=cex.lab, main="Density") 
text(powers,pst$Density,labels=powers,cex=cex1,col="red") 
plot(powers,pst$Heterogeneity,type="n",xlab="Soft Threshold", ylab="Heterogeneity",cex.main=cex.main, cex.lab=cex.lab, cex.axis=cex.axis, main="Heterogeneity") 
text(powers,pst$Heterogeneity,labels=powers,cex=cex1,col="red") 
plot(powers,pst$Centralization,type="n",xlab="Soft Threshold", ylab="Centralization",cex.axis=cex.axis, cex.main=cex.main, cex.lab=cex.lab, main="Centralization") 
text(powers, pst$Centralization, labels=powers,cex=cex1, col="red")
  
```

#Pearson Correlation Adjacency Table
Comparing Adjacency Quantiles between powers 7, 10, 11
Quantiles of the Adjacency Matrix: 
The prob = 0 quantile of a set of numbers is the minumum, the prob = 0.25 quantile is the first quantile, the prob = 0.50 quantile is the median, and prob = 1.0 is the maxium. 
The median adjacency quantile, prob = 0.50, can be considered a robust measure of the network density. 

```{r}
#Pearson is the default 
adjacency_pearson_7 = adjacency(datExpr = combat.adj, type = "signed", power = 7) #R^2 = 0.857

adjacency_pearson_10 = adjacency(datExpr = combat.adj, type = "signed", power = 10) #R^2 = 0.897

adjacency_pearson_11= adjacency(datExpr = combat.adj, type = "signed", power = 11) #R^2 = 0.905
```

```{r}
colQAdj_7 = colQuantileC(adjacency_pearson_7, p = 0.50)
colQAdj_10 = colQuantileC(adjacency_pearson_10, p = 0.50)
colQAdj_11 = colQuantileC(adjacency_pearson_11, p = 0.50)

```




#Building a Network with WGCNA 
```{r}
# here we define the adjacency matrix using soft thresholding with beta=6
ADJ1=abs(cor(combat.adj,use="p"))^11
# When you have a lot of genes use the following code
k=softConnectivity(datE=combat.adj,power=11)
# Plot a histogram of k and a scale free topology plot
sizeGrWindow(10,5)
par(mfrow=c(1,2))
hist(k)
scaleFreePlot(k, main="Check scale free topology\n")
dissTOM=TOMdist(ADJ1)
```
#Module definition using the topological overlap based dissimilarity
```{r}
 #Calculate the dendrogram height 
hierTOM = hclust(as.dist(dissTOM),method="average");
# The reader should vary the height cut-off parameter h1
# (related to the y-axis of dendrogram) in the following
colorStaticTOM = as.character(cutreeStaticColor(hierTOM, cutHeight=.99, minSize=20))
colorDynamicTOM = labels2colors (cutreeDynamic(hierTOM,method="tree"))
colorDynamicHybridTOM = labels2colors(cutreeDynamic(hierTOM, distM= dissTOM , cutHeight = 0.99,
                     deepSplit=2, pamRespectsDendro = FALSE))
# Now we plot the results
sizeGrWindow(10,5)
plotDendroAndColors(hierTOM,
             colors=data.frame(colorStaticTOM,
                              colorDynamicTOM, colorDynamicHybridTOM),
             dendroLabels = FALSE, marAll = c(1, 8, 3, 1),
             main = "Gene dendrogram and module colors, TOM dissimilarity")
```
#Hierarchical clustering dendrogram of module eigengenes.
```{r}
#To get a sense of how related the modules are one can summarize each module by its eigengene (first principal component).
datME=moduleEigengenes(combat.adj, colorDynamicHybridTOM)$eigengenes
signif(cor(datME, use="p"), 2)

#We define a dissimilarity measure between the module eigengenes that keeps track of the sign of the correlation between the module eigengenes, and use it to cluster the eigengene
dissimME=(1-t(cor(datME, method="p")))/2
hclustdatME=hclust(as.dist(dissimME), method="average" )
# Plot the eigengene dendrogram
par(mfrow=c(1,1))
plot(hclustdatME, main="Clustering tree based of the module eigengenes")
```
# The top row shows theatmap of the green module genes (rows) across the microarrays (columns). The lower row shows the corresponding module eigengene expression values (y-axis) versus the same microarray samples. Note that the module eigengene takes on low values in arrays where a lot of module genes are under-expressed (green color in the heatmap). The ME takes on high values for arrays where a lot of module genes are over-expressed (red in the heatmap). ME can be considered the most representative gene expression profile of the module.

#ME = module eigengenes

#minAbsSplitHeigh = 0.35 in cutreeDynamic
#eigengene and hub genes
```{r}
sizeGrWindow(8,7);
which.module="turquoise"
ME=datME[, paste("ME",which.module, sep="")]
par(mfrow=c(2,1), mar=c(0.3, 5.5, 3, 2))
plotMat(t(scale(combat.adj[ ,colorDynamicHybridTOM==which.module ]) ),
        nrgcols=30,rlabels=F,rcols=which.module,
        main=which.module, cex.main=2)
par(mar=c(5, 4.2, 0, 0.7))
barplot(ME, col=which.module, main="", cex.main=2,
       ylab="eigengene expression",xlab="array sample")
```

#BiWeight Correlation - WGCNA
Didn't supply me with a power estimate approaching 0.90 
```{r}
powers = 1:20
#Call the network topology analysis function
#Analysis of scale free topology for multiple hard thresholds
sft_bicor = pickSoftThreshold(data = combat.adj, RsquaredCut = 0.90, dataIsExpr = T, powerVector = powers,networkType ="signed", corFnc="bicor" )
# Plot the results:
sizeGrWindow(9, 5)
par(mfrow = c(1,2));
cex1 = 0.9;
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft_bicor$fitIndices[,1], -sign(sft_bicor$fitIndices[,3])*sft_bicor$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"));
text(sft_bicor$fitIndices[,1], -sign(sft_bicor$fitIndices[,3])*sft_bicor$fitIndices[,2],
     labels=powers,cex=cex1,col="red");
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft_bicor$fitIndices[,1], sft_bicor$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft_bicor$fitIndices[,1], sft_bicor$fitIndices[,5], labels=powers, cex=cex1,col="red")

```

#Spearman Correlation
```{r}
powers = 1:20
#Call the network topology analysis function
#Analysis of scale free topology for multiple hard thresholds
sft = pickSoftThreshold(data = combat.adj, RsquaredCut = 0.90, dataIsExpr = T, powerVector = powers,networkType ="signed", corFnc="cor", corOptions = list(method = "spearman"))
# Plot the results:
sizeGrWindow(9, 5)
par(mfrow = c(1,2));
cex1 = 0.9;
# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red");
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```
#KEGG biological pathway interpretation for modules 
#EnrichR 
