---
title: "PFAS_eigengene_analysis"
output: html_document
date: "2023-06-8"
---
```{r}
library(tidyverse)
library(gprofiler2)
library(dplyr)
load("combat.adj.Rdata")
load("eigengnes_spearman_9.Rdata")
load("log_PFAS_pheno.Rdata")

#("G-155","G-157") are not in the pfas data 
combat.adj = t(combat.adj)
combat.adj = combat.adj[!row.names(combat.adj) %in% c("G-155","G-157"),]
eigengenes = module_Eigengenes$eigengenes
eigengenes = eigengenes[!row.names(eigengenes) %in% c("G-155","G-157"),]

pc <- prcomp(combat.adj)
pc_data = pc$x
pc_data = pc_data[,"PC1"]


PFASp = c("L-PFOA","PFHxS","PFNA","PFDA","s-PFOS")
log_pfas = log_pfas[rownames(eigengenes),PFASp]
pheno_139 = pheno[rownames(eigengenes),]

#pheno_139 = pheno_139[-37,]
#pc_data = pc_data[-37]
#log_pfas = log_pfas[-37,]
#eigengenes = eigengenes[-37,]

```


```{r}
xist_sex = data.frame(sex = pheno_139$childs_sex, XIST = combat.adj[,"XIST"], PFNA = log_pfas$PFNA, ID = pheno_139$participant_id)

ggplot(data = xist_sex, aes(x = PFNA, y = XIST, color = sex, )) + geom_point() + geom_text(aes(label=ifelse(sex == "Male" & XIST >12 ,as.character(ID),'')),hjust=0,vjust=0)
```



```{r}
library(reshape2)
cor_pfas = cor(log_pfas)

cor_melt <- melt(cor_pfas)



ggplot(cor_melt, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), size = 4) +  # Adding correlation values with label size = 10
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal() +
  labs(x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 12),  # Adjust x-axis text size
        axis.text.y = element_text(size = 12), legend.text = element_text(size = 10))  # Adjust y-axis text size

```

```{r}
library(corrplot)

corrplot(cor_pfas, method = 'number', order = 'FPC', type = 'upper', diag = FALSE)
```


#eigengenes and pfas 
```{r}
eigengene_PFAS <- data.frame()


for (i in 1:ncol(eigengenes)) {
  for (x in PFASp){
    model <- lm(eigengenes[, i] ~ log_pfas[, x] + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks +  pheno_139$childs_sex + pheno_139$MomBMI_10wks + pheno_139$official_enroll_category+ pc_data)
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, module = colnames(eigengenes)[i], PFAS = x)
    eigengene_PFAS <- rbind(eigengene_PFAS, df)
  }
}


eigengene_PFAS$FDR = p.adjust(eigengene_PFAS$p.value, method = "fdr", n =105)

 eigengene_PFAS %>%
  filter(FDR <= 0.05) %>%
  mutate(module = gsub("ME", "", module))
 
```
#fdr of eigengene ~PFAS linear regression models 
```{r}

```



#birth lenght, birth weight, and GA as outcomes (module eigengenes not included in model)
```{r}
birthlength = data.frame()
for (x in PFASp){
     model <- lm( pheno_139$birth_len_cm ~ log_pfas[,x] + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks +  pheno_139$childs_sex + pc_data)
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, outcome = "Birth Length" , PFAS = x)
    birthlength <- rbind(birthlength, df)
}

birthlength %>% filter(p.value <= 0.05)

birthweight = data.frame()
for (x in PFASp){
     model <- lm( pheno_139$birth_wt_kg ~ log_pfas[, x] + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks +  pheno_139$childs_sex + pc_data)
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, outcome = "Birthweight" , PFAS = x)
    birthweight <- rbind(birthweight, df)
}

birthweight %>% filter(p.value <= 0.05)

GA = data.frame()
for (x in PFASp){
     model <- lm(pheno_139$gest_age_in_weeks_edd ~ log_pfas[, x] + pheno_139$mom_age_at_birth + pheno_139$mom_education +  pheno_139$MomBMI_36wks +  pheno_139$childs_sex + pc_data)
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, outcome = "GA" , PFAS = x)
    GA <- rbind(GA, df)
}

GA %>% filter(p.value <= 0.05)
```

```{r}
birthlength = data.frame()
for (x in PFASp){
  for (i in 1:ncol(eigengenes)) {
     model <- lm( pheno_139$birth_len_cm ~ log_pfas[, x] + eigengenes[,i]+ pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks +  pheno_139$childs_sex + pc_data)
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, outcome = "Birth Length" , PFAS = x, module = colnames(eigengenes)[i])
    birthlength <- rbind(birthlength, df)
  }
}

birthlength %>% filter(p.value <= 0.05)

birthweight = data.frame()
for (x in PFASp){
  for (i in 1:ncol(eigengenes)) {
     model <- lm( pheno_139$birth_wt_kg ~ log_pfas[, x] + eigengenes[,i] + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks +  pheno_139$childs_sex + pc_data)
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, outcome = "Birthweight" , PFAS = x, module =colnames(eigengenes)[i])
    birthweight <- rbind(birthweight, df)
  }}

birthweight %>% filter(p.value <= 0.05)

GA = data.frame()
for (x in PFASp){
  for (i in 1:ncol(eigengenes)) {
     model <- lm(pheno_139$gest_age_in_weeks_edd ~ log_pfas[, x] + eigengenes[,i] + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$MomBMI_36wks +  pheno_139$childs_sex + pc_data)
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, outcome = "GA" , PFAS = x, module =colnames(eigengenes)[i])
    GA <- rbind(GA, df)
}}

GA %>% filter(p.value <= 0.05) %>% mutate(module = gsub("ME", "", module))
```

#dataframes for sex stratification created used first for GA 
```{r}
f_pheno = pheno_139[which(pheno_139$childs_sex == "Female"),]

m_pheno = pheno_139[which(pheno_139$childs_sex == "Male"),]

pfas_f = log_pfas[which(pheno_139$childs_sex == "Female"),]
  
pfas_m = log_pfas[which(pheno_139$childs_sex == "Male"),]

eigengene_m = eigengenes[which(pheno_139$childs_sex == "Male"),]
eigengene_f = eigengenes[which(pheno_139$childs_sex == "Female"),]

GA_f = data.frame()
for (x in PFASp){
  for (i in 1:ncol(eigengenes)) {
     model <- lm(f_pheno$gest_age_in_weeks_edd ~ pfas_f[, x] + eigengene_f[,i] + f_pheno$mom_age_at_birth + f_pheno$mom_education + f_pheno$MomBMI_36wks + pc_data[which(pheno_139$childs_sex == "Female")])
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, outcome = "GA" , PFAS = x, module =colnames(eigengene_f)[i])
    GA_f <- rbind(GA_f, df)
}}

GA_f %>% filter(p.value <= 0.05) %>% mutate(module = gsub("ME", "", module))

GA_m = data.frame()
for (x in PFASp){
  for (i in 1:ncol(eigengenes)) {
     model <- lm(m_pheno$gest_age_in_weeks_edd ~ pfas_m[, x] + eigengene_m[,i] + m_pheno$mom_age_at_birth + m_pheno$mom_education + m_pheno$MomBMI_36wks + pc_data[which(pheno_139$childs_sex == "Male")])
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, outcome = "GA" , PFAS = x, module =colnames(eigengene_m)[i])
    GA_m <- rbind(GA_m, df)
}}

GA_m %>% filter(p.value <= 0.05) %>% mutate(module = gsub("ME", "", module))


```

#birth weight stratified by sex 
```{r}
BW_f = data.frame()
for (x in PFASp){
  for (i in 1:ncol(eigengenes)) {
     model <- lm(f_pheno$birth_wt_kg ~ pfas_f[, x] + eigengene_f[,i] + f_pheno$gest_age_in_weeks_edd + f_pheno$mom_age_at_birth + f_pheno$mom_education + f_pheno$MomBMI_36wks + pc_data[which(pheno_139$childs_sex == "Female")])
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
df <- data.frame(p.value = p_value, estimate = estimate, outcome = "BW" , PFAS = x, module=colnames(eigengene_f)[i])
BW_f <- rbind(BW_f, df)
  }
    }

BW_f %>% filter(p.value <= 0.05) %>% mutate(module = gsub("ME", "", module))

BW_m = data.frame()
for (x in PFASp){
  for (i in 1:ncol(eigengenes)) {
     model <- lm(m_pheno$birth_wt_kg ~ pfas_m[, x] + eigengene_m[,i] + m_pheno$gest_age_in_weeks_edd + m_pheno$mom_age_at_birth + m_pheno$mom_education + m_pheno$MomBMI_36wks + pc_data[which(pheno_139$childs_sex == "Male")])
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, outcome = "BW" , PFAS = x, module =colnames(eigengene_m)[i])
    BW_m <- rbind(BW_m, df)
    
  }
  }

BW_m %>% filter(p.value <= 0.05) %>% mutate(module = gsub("ME", "", module))
```

#BL regression analysis stratified by sex
```{r}
BL_f = data.frame()
for (x in PFASp){
  for (i in 1:ncol(eigengenes)) {
     model <- lm(f_pheno$birth_len_cm ~ pfas_f[, x] + eigengene_f[,i] + f_pheno$gest_age_in_weeks_edd + f_pheno$mom_age_at_birth + f_pheno$mom_education + f_pheno$MomBMI_36wks + pc_data[which(pheno_139$childs_sex == "Female")])
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, outcome = "BL" , PFAS = x, module=colnames(eigengene_f)[i])
    BL_f <- rbind(BL_f, df)
}}

BL_f %>% filter(p.value <= 0.05)

BL_m = data.frame()
for (x in PFASp){
  for (i in 1:ncol(eigengenes)) {
     model <- lm(m_pheno$birth_len_cm ~ pfas_m[, x] + eigengene_m[,i] + m_pheno$gest_age_in_weeks_edd + m_pheno$mom_age_at_birth + m_pheno$mom_education + m_pheno$MomBMI_36wks + pc_data[which(pheno_139$childs_sex == "Male")])
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, outcome = "BL" , PFAS = x, module =colnames(eigengene_m)[i])
    BL_m <- rbind(BL_m, df)
}}

BL_m %>% filter(p.value <= 0.05)
```

#quantile regression 
```{r}
library(quantreg)

df_50 = data.frame()
for( i in 1:ncol(eigengenes)){
  model <- rq(eigengenes[,i] ~ log_pfas[,"PFDA"] + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks + pheno_139$childs_sex + pc_data, tau = 0.50)
  rq_summary = summary(model, se= "nid")
  p_value <- rq_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, module=colnames(eigengenes)[i])
   df_50 <- rbind(df_50, df)
}

df_50 = df_50 %>% filter(p.value <= 0.05) %>% mutate(module = gsub("ME", "", module))

df_75 = data.frame()
for( i in 1:ncol(eigengenes)){
  model <- rq(eigengenes[,i] ~ log_pfas[,"PFDA"] + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks + pheno_139$childs_sex + pc_data, tau = 0.75)
  rq_summary = summary(model, se= "nid")
   p_value <- rq_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, module=colnames(eigengenes)[i])
   df_75 <- rbind(df_75, df)
}

df_75 = df_75 %>% filter(p.value <= 0.05) %>% mutate(module = gsub("ME", "", module))
```

#eigengenes~PFAS model with interaction between pfas and sex 
```{r}
eigengene_PFAS_sex <- data.frame()

for (i in 1:ncol(eigengenes)){
  for (x in PFASp){
    model <- lm(eigengenes[,i] ~ log_pfas[,x] + pheno_139$childs_sex + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks + + pheno_139$MomBMI_10wks + pheno_139$official_enroll_category + pc_data + pheno_139$childs_sex:log_pfas[,x])
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[12,4]
    df <- data.frame(p.value = p_value, module = colnames(eigengenes)[i], PFAS = x)
    eigengene_PFAS_sex <- rbind(eigengene_PFAS_sex, df)
 }
}

#eigengene_PFAS_sex %>% filter(.$p.value <= 0.05)

eigengene_PFAS_sex$p.adjust = p.adjust(eigengene_PFAS_sex$p.value, method = "fdr", n =105)

eigengene_PFAS_sex %>% filter(.$p.adjust <= 0.05)

```

#enrichment analysis with TFN 
```{r}
TF_enrichment = data.frame()
load("DataFile_FinalTRN.RData")

#unique(eigengene_PFAS$module)
for ( i in "darkred"){
  color = gene_color %>% filter(color == i) %>% pull(gene_name)
  TFN = TRNModel %>% filter(.$targetGene %in% color) 
  sorted_df = sort(table(TFN$TF), decreasing = TRUE)
  sorted_df = as.data.frame(sorted_df)
  colnames(sorted_df) = c("Variable", "Count")
  
  in_module = sorted_df[1, "Variable"] %in% color
  
  ppars = sorted_df %>% filter(grepl("PPAR", Variable)) 
  if(nrow(ppars) == 0){
     df= data.frame(top_TF = sorted_df[1, "Variable"], PPARA = 0, PPARD = 0, PPARG = 0, top_TF_module= in_module,  top_TF_num = sorted_df[1, "Count"], module = i)
  } else {
    rownames(ppars) = ppars$Variable
    df= data.frame(top_TF = sorted_df[1, "Variable"], PPARA = ppars["PPARA","Count"], PPARD = ppars["PPARD","Count"], PPARG =ppars["PPARG","Count"], top_TF_module= in_module, top_TF_num = sorted_df[1, "Count"], module = i)
  }
  
TF_enrichment  = rbind(TF_enrichment, df)
}


TF_enrichment[is.na(TF_enrichment)] <- 0

TF_enrichment
```

# analysis of dark turquoise (violin plots)
```{r}
library(biomaRt)
library(cowplot)

ensembl=useMart("ensembl")


ensembl=useDataset("hsapiens_gene_ensembl",mart = ensembl)
ensembl.ids = colnames(combat.adj)

# hgnc_symbol = refseq gene name?
#everywhere with ensembl_gene_id change to hgnc_symbol
ensembl.genes <- getBM(attributes = c('hgnc_symbol', 'chromosome_name','start_position', 'end_position'),
                       filters = 'hgnc_symbol', 
                       values = ensembl.ids, 
                       mart = ensembl)

darkturquoise = gene_color %>% filter(color == "darkturquoise") %>% pull(gene_name)

darkturquoise_gene_position = ensembl.genes %>% filter(.$hgnc_symbol %in% darkturquoise) 
#all of them in x and y chromosomes only 21 genes shown not all 28 
 

#can produce a violin plot of expression by sex for the genes within dark turqouise 
violin_plot_turq = combat.adj[, darkturquoise]
violin_plot_turq = data.frame(violin_plot_turq)
violin_plot_turq <- rownames_to_column(violin_plot_turq, "participant_id")
col_to_merge = pheno_139[,c("participant_id","childs_sex")]

violin_plot_turq = merge(violin_plot_turq,col_to_merge, by = "participant_id")

plots <- list()

# Iterate over the darkturquoise list and create violin plots
for (variable in darkturquoise) {
  plot <- ggplot(data = violin_plot_turq, aes(x = childs_sex, y = !!rlang::sym(variable))) +
    geom_violin() +
    labs(x = NULL, y = variable)  # Set the y-axis label as the variable name
  
  plots[[length(plots) + 1]] <- plot  # Append the plot to the list
}

plot_grid(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]], plots[[10]], plots[[11]], plots[[12]], plots[[13]], plots[[14]])


plot_grid(plots[[15]], plots[[16]], plots[[17]], plots[[18]], plots[[19]], plots[[20]], plots[[21]], plots[[22]], plots[[23]], plots[[24]], plots[[25]], plots[[26]], plots[[27]], plots[[28]])

```
#PFAS and PPAR expression 
```{r}
PPAR_pfas = data.frame()
for( x in PFASp){
  for ( i in c("PPARA", "PPARD","PPARG")){
    model <- lm(combat.adj[, i] ~ log_pfas[, x] + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks + pheno_139$childs_ethnicity + pheno_139$childs_sex + pc_data)
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, PPAR = i, PFAS = x)
     PPAR_pfas <- rbind(PPAR_pfas, df)
  }
}

PPAR_pfas = PPAR_pfas %>% filter(p.value <= 0.05)

PPAR_pfas
```
#dark turquoise~PFAS linear regression model with sex interaction 
```{r}
 dark_red = data.frame()
for (x in PFASp){
  model2 <- summary(lm(eigengenes$MEdarkred ~ log_pfas[,x] + pheno_139$childs_sex + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks + pheno_139$childs_ethnicity + pheno_139$childs_sex + pc_data + log_pfas[,x]:pheno_139$childs_sex))
  
  model1 <- summary(lm(eigengenes$MEdarkred ~ log_pfas[,x] + pheno_139$childs_sex))
  
  #summary for one model at a time pull the coefficients at a time 
  df = data.frame(p_value_no_inter = model1$coefficients[2,4], p_value_inter = model2$coefficients[11,4], r2_no_inter = model1$r.squared, r2_inter = model2$r.squared, no_inter_estimate = model1$coefficients[1,3], inter_estimate = model2$coefficients[1,3], PFAS =x)
  
  dark_red  = rbind(dark_red, df)
}
dark_red %>% filter(p_value_inter <= 0.05)
```

#pink KEGG and GO Terms 
```{r}
library(gprofiler2)
pink = gene_color %>% filter(color == "darkred") %>% pull(gene_name)
#pink_KEGG = gost(pink, organism = "hsapiens", sources = "KEGG")$result

#ggplot(data = pink_KEGG, aes(x = -log10(p_value), y = term_name )) + geom_point(aes(size = intersection_size, color = factor(source))) + ylab("") + xlab(expression("-log"[10]*"(p-value)")) + labs(size="Hits") + ggtitle("Dark Red KEGG Terms") + theme(plot.title = element_text(face = "bold", size = 20, hjust = 0.5), axis.text.x= element_text(size = 10), axis.text.y= element_text(size = 10))+ scale_colour_brewer(palette = "Set2")

pink_GO = gost(pink, organism = "hsapiens", sources = "GO")$result

ggplot(data = pink_GO, aes(x = -log10(p_value), y = term_name )) + geom_point(aes(size =intersection_size , color = factor(source))) + ylab("") + xlab(expression("-log"[10]*"(p-value)")) + labs(size="Hits", color="GO source") + ggtitle("Dark Red") + theme(plot.title = element_text(face = "bold", size = 20, hjust = 0.5), axis.text.x= element_text(size = 10), axis.text.y= element_text(size = 5))+ scale_colour_brewer(palette = "Set2")
```
#green yellow KEGG and GO 
```{r}
greenyellow = gene_color %>% filter(color == "darkred") %>% pull(gene_name)
greenyellow_KEGG = gost(greenyellow, organism = "hsapiens", sources = "KEGG")$result

ggplot(data = greenyellow_KEGG, aes(x = -log10(p_value), y = term_name )) + geom_point(aes(size = intersection_size, color = factor(source))) + ylab("") + xlab(expression("-log"[10]*"(p-value)")) + labs(size="Hits") + ggtitle("Dark Red KEGG Terms") + theme(plot.title = element_text(face = "bold", size = 20, hjust = 0.5), axis.text.x= element_text(size = 10), axis.text.y= element_text(size = 10))+ scale_colour_brewer(palette = "Set2")

greenyellow_GO = gost(greenyellow, organism = "hsapiens", sources = "GO")$result

ggplot(data = greenyellow_GO, aes(x = -log10(p_value), y = term_name )) + geom_point(aes(size = intersection_size, color = factor(source))) + ylab("") + xlab(expression("-log"[10]*"(p-value)")) + labs(size="Hits", color="GO source") + ggtitle("Dark Red GO Terms") + theme(plot.title = element_text(face = "bold", size = 20, hjust = 0.5), axis.text.x= element_text(size = 10), axis.text.y= element_text(size = 10))+ scale_colour_brewer(palette = "Set2")
```
#hypeR gene enrichment of the molecular signatures collection of  “hallmark” genes
```{r}
#devtools::install_github("montilab/hypeR")
library(hypeR)
colors <- unique(gene_color$color)

# Initialize an empty list to store the result
signatures <- list()

# Iterate over each color
for (color in colors) {
  # Filter genes for the current color
  genes <- gene_color$gene_name[gene_color$color == color]
  
  # Add color and corresponding genes to the result list
  signatures[[color]] <- genes
}

signatures = signatures[names(signatures) != "grey"] 

genesets <- msigdb_gsets("Homo sapiens", "H", "", clean=TRUE)

mhyp <- hypeR(signatures, genesets, test="hypergeometric", background=30000)

hyp_dots(mhyp, merge=TRUE, fdr=0.05, title="Co-expression Modules")
```

#XY module/dark red  stratified by sex 
```{r}
f_pheno = pheno_139[which(pheno_139$childs_sex == "Female"),]

m_pheno = pheno_139[which(pheno_139$childs_sex == "Male"),]

pfas_f = log_pfas[which(pheno_139$childs_sex == "Female"),]
  
pfas_m = log_pfas[which(pheno_139$childs_sex == "Male"),]

eigengene_m = eigengenes[which(pheno_139$childs_sex == "Male"),]
eigengene_f = eigengenes[which(pheno_139$childs_sex == "Female"),]


eigengene_PFAS_m<- data.frame()


  for (x in colnames(log_pfas)){
    model <- lm(eigengene_m[, "MEdarkred"] ~ log_pfas[rownames(m_pheno), x] + m_pheno$mom_age_at_birth + m_pheno$mom_education + m_pheno$gest_age_in_weeks_edd + m_pheno$MomBMI_36wks +  pc_data[which(pheno_139$childs_sex == "Male")])
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, module = "Dark Red", PFAS = x)
    eigengene_PFAS_m <- rbind(eigengene_PFAS_m, df)
  }


 eigengene_PFAS_m 

 eigengene_PFAS_m
 
 
 eigengene_PFAS_f<- data.frame()


  for (x in colnames(log_pfas)){
    model <- lm(eigengene_f[, "MEdarkred"] ~ log_pfas[rownames(f_pheno), x] + f_pheno$mom_age_at_birth + f_pheno$mom_education + f_pheno$gest_age_in_weeks_edd + f_pheno$MomBMI_36wks +  pc_data[which(pheno_139$childs_sex == "Female")])
    lm_summary <- summary(model)
    p_value <- lm_summary$coefficients[2,4]
    estimate <- coef(model)[2]
    df <- data.frame(p.value = p_value, estimate = estimate, module = "Dark Red", PFAS = x)
    eigengene_PFAS_f <- rbind(eigengene_PFAS_f, df)
  }


 


 eigengene_PFAS_f 

 
 
 
```
#quatiles into categorical variables
```{r}

# Compute quantiles
quantiles <- quantile(log_pfas$PFDA, probs = seq(0, 1, by = 0.25))

if (quantiles[1] == min(log_pfas$PFDA)) {
  quantiles[1] <- quantiles[1] - 1e-10
}

# Create quantile column
quantile_column <- cut(log_pfas$PFDA, breaks = quantiles, labels = FALSE, include.lowest = TRUE)

quantile_column = as.character(quantile_column)
eigengene_PFDA_quantile <- data.frame()

for (i in 1:ncol(eigengenes)) {
    model <- lm(eigengenes[, i] ~ quantile_column + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks +  pheno_139$childs_sex + pheno_139$MomBMI_10wks + pheno_139$official_enroll_category+  pc_data)
    lm_summary <- summary(model)
    df <- data.frame(  p_value_Q2 =lm_summary$coefficients[2,4], estimate_Q2 = lm_summary$coefficients[2,1], p_value_Q3 = lm_summary$coefficients[3,4], estimate_Q3 = lm_summary$coefficients[3,1],  p_value_Q4 = lm_summary$coefficients[4,4], estimate_Q4 = lm_summary$coefficients[4,1], module = colnames(eigengenes)[i], PFAS = "PFDA")
   eigengene_PFDA_quantile <- rbind(eigengene_PFDA_quantile, df)
  }


eigengene_PFDA_quantile = eigengene_PFDA_quantile %>%
  filter(p_value_Q2 <= 0.05 | p_value_Q3 <= 0.05 |p_value_Q4 <= 0.05 ) %>%
  mutate(module = gsub("ME", "", module))
 
 eigengene_PFDA_quantile
 
 table(quantile_column)

```

```{r}
 model <- lm(eigengenes[,"MEblack"] ~ quantile_column + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks +  pheno_139$childs_sex + pc_data)
    lm_summary <- summary(model)
    
     lm_summary
```

```{r}
 model <- lm(eigengenes[, i] ~ log_pfas[, x] + pheno_139$mom_age_at_birth + pheno_139$mom_education + pheno_139$gest_age_in_weeks_edd + pheno_139$MomBMI_36wks +  pheno_139$childs_sex + pc_data)


```

```{r}
library(openxlsx)

# Assuming you have multiple data frames named eigengene_PFAS, df_75, df_50, eigengene_PFDA_quantile, eigengene_PFAS_sex, ...
# and you want to export them to different sheets in the same Excel file.

# Define the file path for the Excel file
excel_file <- "/Users/cynthiaperez/WGCNA_PFAS/Aug7.xlsx"

# Create a new workbook
wb <- createWorkbook()

# Add each data frame to a separate sheet
addWorksheet(wb, sheetName = "Sheet1")
writeData(wb, sheet = "Sheet1", eigengene_PFAS)


addWorksheet(wb, sheetName = "Quant_Cat")
writeData(wb, sheet = "Quant_Cat", eigengene_PFDA_quantile)

addWorksheet(wb, sheetName = "sex_PFAS")
writeData(wb, sheet = "sex_PFAS", eigengene_PFAS_sex)

# Save the workbook to the Excel file
saveWorkbook(wb, excel_file)
```

